\bf Nonparametric Bootstrap Kolmogorov-Smirnov Test for Goodness-of-Fit of Marginal Distributions of Stationary Time Series

### 2023-12-05 manuscript/manuscript.tex
meta cleaning in preamble
comments and minor edits

% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

@@

\if0\blind
{
  \title{\bf Nonparametric Bootstrap Kolmogorov-Smirnov Test with Stationary 
  Time Series}
  \author{Mathew Chandy, %\\
%   \href{mailto:mathew.chandy@uconn.edu}
% {\nolinkurl{mathew.chandy@uconn.edu}}\\
  Jun Yan, %\\
  Xianyang Zhang\\
}
\date{}
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Nonparametric Bootstrap Kolmogorv-Smirnov Test with Stationary 
    Time Series}
\end{center}
  \bigskip
} \fi

@@

%\doublespace

@@

@@

Consider a stationary time series $\{X_t: t = 1, \ldots, n\}$ with length~$n$.
\jy{state the null hypothesis explicitly; we are testing the marginal
  distribution. Point out the challenges.}
Let $H_0: X_t ~ F$, where if $F$ is a hypothesized distribution with unknown 
parameters. Let $H_A$ be that $X_t$ does not follow $F$. This is challenging because we are
dealing with a situation with both unknown parameters and temporal dependence.
If $X_t ~ F$, the true marginal distribution of $X_t$ has some parameters 
$\theta$.

\jy{define $\hat\theta_n$}
\mc{addressed}

\jy{No double letter notations}
\mc{addressed}
Consider the goodness of fit statistic:
\begin{equation*}
  T_n := \sup_x|Y_n(x; \hat\theta)|, 
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n)),
\end{equation*}
where $F_n$ denotes the empirical distribution function based on $X_1,...,X_n$,
and $\hat\theta_n$ denotes the fitted parameters for the hypothesized 
distribution fitted onto $X_t$.
We note that
\begin{equation*}
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x)) - 
\sqrt{n}(F(x; \hat\theta_n) - F(x)),
\end{equation*}
where $F(x)$ is the true cdf (under the null $F(x) = F(x, \theta_0)$ for some
true parameter $\theta_0$).

Let us first consider the case where $X_i$'s are i.i.d. Denote by $F^{(b)}_n$ the
empirical distribution of the $b$th bootstrap sample and let
$\hat\theta^{(b)}_n$ be the parameter estimate based on the bootstrap samples. 
Using the bootstrap (asymptotic) theory, we can approximate the distribution of
\jy{no need for so many in-display equations}
$\sqrt{n}(F_n(x) - F(x))$
by that of $\sqrt{n}(F^{(b)}_n(x) - F_n(x))$
and the distribution of
$\sqrt{n}(F(x; \hat\theta_n) - F(x))$
by that of
$\sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; \hat\theta_n))$.
Therefore, if we define
\begin{equation*}
Y^{(b)}_n(x) = \sqrt{n}(F^{(b)}_n(x) - F_n(x)) - 
\sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; \hat\theta_n)) 
= \sqrt{n}(F^{(b)}_n(x) - F(x; \hat\theta^{(b)}_n)) - 
\sqrt{n}(F_n(x) - F(x; \hat\theta_n)),
\end{equation*}
then $T^{(b)}_n := \sup_x|Y^{(b)}_n(x)|$ is the bootstrap statistic that is expected
to approximate the distribution of $T_n$. We note that the term
$\sqrt{n}(F_n(x) - F(x; \hat\theta_n))$ is exactly the bias term considered in 
Babu and Rao (2004).

We now consider the case where $X_i$'s are realizations from a time series and
$X^{(b)}_1,...,X^{(b)}_n$ are generated by block bootstrap for 
$b \in \{1, \ldots, B\}$. Let $E^{(b)}$ refer to the expectation with respect to
the bootstrap distribution
In this case, we can 
approximate the distribution of
$\sqrt{n}(F_n(x) - F(x))$
by that of
$\sqrt{n}(F^{(b)}_n(x) - E^{(b)}[F^{(b)}_n(x)])$,
where 
\jy{change the languages to add variety}
and the distribution of $\sqrt{n}(F(x; \theta_n) - F(x))$
by that of $\sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; E^*[\hat\theta^{(b)}_n]))$.
\jy{no closing a sentence before all the notations are introduced}
We can compute $E^{(b)}[F^{(b)}_n(x)]$ and 
$E^{(b)}[\hat\theta^{(b)}_n]$ numerically (they can also be computed analytically, 
depending on the types of block bootstrap we use). One can compute 
$F^{(b)}_n$ 
and $\hat\theta^{(b)}_n$ based on
the $b$th bootstrap sample. Then
$E^{(b)}[F^{(b)}_n(x)] \approx \frac{1}{B}\sum_{b = 1}^BF^{(b)}_n(x)$, and
$E^{(b)}[\hat\theta^{(b)}_n] \approx \frac{1}{B}\sum_{b = 1}^B\hat\theta^{(b)}_n$.
In this case, we can define
\jy{too long}

$Y^*_n(x) = \sqrt{n}(F^{(b)}_n(x) - E^{(b)}[F^{(b)}_n(x)]) - 
\sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; E^{(b)}[\hat\theta^{(b)}_n)]) \\
= \sqrt{n}(F^{(b)}_n(x) - F(x; \hat\theta^{(b)}_n)) - 
\sqrt{n}(E^{(b)}[F^{(b)}_n(x)] - F(x; E^{(b)}[\hat\theta^{(b)}_n]))$,

and $T^{(b)}_n = \sup_x|Y^{(b)}_n(x)|$.

The p-value of the block bootstrap KS test can be computed 
as $p = \#\{T^{(b)}_n < T_n\} / B$ for 
$b \in \{1, \ldots, B\}$.

@@

\jy{First size and then power}

>>>>

% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0]}

@@

\title{\bf Nonparametric Bootstrap Kolmogorov-Smirnov Test for
  Goodness-of-Fit of Marginal Distributions of Stationary Time Series}
\if0\blind
{
  \author{Mathew Chandy, %\\
%   \href{mailto:mathew.chandy@uconn.edu}
%   {\nolinkurl{mathew.chandy@uconn.edu}}\\
  Elizabeth Schifano\\
  Jun Yan, %\\
  Xianyang Zhang\\
}
} \fi

\maketitle

@@

@@

\jy{Flow:
  Review misuses of one-sample KS test; focus on serial dependence;
  criticize Zeimbakakis; highlight contributions (connection to Babu and Rao).}

\jy{Borrow the introduction of Zeimbakakis and, of course, rephase them.}

@@

\jy{Clarify the GoF test is for the marginal distribution of a statoinary
  series.}
Consider a stationary time series $\{X_t: t = 1, \ldots, n\}$ with length~$n$.
\jy{state the null hypothesis explicitly; we are testing the marginal
  distribution. Point out the challenges.}
Let $H_0: X_t ~ F$, where if $F$ is a hypothesized distribution with unknown 
parameters. Let $H_A$ be that $X_t$ does not follow $F$. This is challenging because we are
dealing with a situation with both unknown parameters and temporal dependence.
If $X_t ~ F$, the true marginal distribution of $X_t$ has some parameters 
$\theta$.

\jy{No double letter notations}
\mc{addressed}
Consider the goodness of fit statistic:
\jy{define $Y_n$ before introducing $T_n$}
\begin{equation*}
  T_n := \sup_x|Y_n(x; \hat\theta)|, 
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n)),
\end{equation*}
where $F_n$ denotes the empirical distribution function based on $X_1,...,X_n$,
and $\hat\theta_n$ denotes the fitted parameters for the hypothesized 
distribution fitted onto $X_t$.
We note that
\begin{equation*}
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x)) - 
\sqrt{n}(F(x; \hat\theta_n) - F(x)),
\end{equation*}
where $F(x)$ is the true cdf (under the null $F(x) = F(x, \theta_0)$ for some
true parameter $\theta_0$).

\jy{respond to my comments}

Let us first consider the case where $X_i$'s are i.i.d. Denote by $F^{(b)}_n$ the
empirical distribution of the $b$th bootstrap sample and let
$\hat\theta^{(b)}_n$ be the parameter estimate based on the bootstrap samples. 
Using the bootstrap (asymptotic) theory, we can approximate the distribution of
\jy{Need rephrasing. The two terms are approximated altogether jointly not
  separately.}
$\sqrt{n}(F_n(x) - F(x))$
by that of $\sqrt{n}(F^{(b)}_n(x) - F_n(x))$
and the distribution of
$\sqrt{n}(F(x; \hat\theta_n) - F(x))$
by that of
$\sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; \hat\theta_n))$.
Therefore, if we define
\begin{equation*}
Y^{(b)}_n(x) = \sqrt{n}(F^{(b)}_n(x) - F_n(x)) - 
\sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; \hat\theta_n)) 
= \sqrt{n}(F^{(b)}_n(x) - F(x; \hat\theta^{(b)}_n)) - 
\sqrt{n}(F_n(x) - F(x; \hat\theta_n)),
\end{equation*}
then $T^{(b)}_n := \sup_x|Y^{(b)}_n(x)|$ is the bootstrap statistic that is expected
to approximate the distribution of $T_n$. We note that the term
$\sqrt{n}(F_n(x) - F(x; \hat\theta_n))$ is exactly the bias term considered in 
Babu and Rao (2004).\jy{cite}

We now consider the case where $X_i$'s are realizations from a time series and
$X^{(b)}_1,...,X^{(b)}_n$ are generated by block bootstrap for 
$b \in \{1, \ldots, B\}$. Let $E^{(b)}$ refer to the expectation with respect to
the bootstrap distribution. In this case, we can 
approximate the distribution of
$\sqrt{n}(F_n(x) - F(x))$
by that of
$\sqrt{n}(F^{(b)}_n(x) - E^{(b)}[F^{(b)}_n(x)])$,
\jy{change the languages to add variety}
and the distribution of $\sqrt{n}(F(x; \theta_n) - F(x))$
by that of $\sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; E^*[\hat\theta^{(b)}_n]))$.

We can compute $E^{(b)}[F^{(b)}_n(x)]$ and 
$E^{(b)}[\hat\theta^{(b)}_n]$ numerically (they can also be computed analytically, 
depending on the types of block bootstrap we use). One can compute 
$F^{(b)}_n$ 
and $\hat\theta^{(b)}_n$ based on
the $b$th bootstrap sample. Then
$E^{(b)}[F^{(b)}_n(x)] \approx \frac{1}{B}\sum_{b = 1}^BF^{(b)}_n(x)$, and
$E^{(b)}[\hat\theta^{(b)}_n] \approx \frac{1}{B}\sum_{b = 1}^B\hat\theta^{(b)}_n$.
In this case, we can define
\begin{align*}
  Y^*_n(x) &= \sqrt{n}(F^{(b)}_n(x) - E^{(b)}[F^{(b)}_n(x)]) - 
             \sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; E^{(b)}[\hat\theta^{(b)}_n)]) \\
           &= \sqrt{n}(F^{(b)}_n(x) - F(x; \hat\theta^{(b)}_n)) -
             \sqrt{n}(E^{(b)}[F^{(b)}_n(x)] - F(x; E^{(b)}[\hat\theta^{(b)}_n])),
\end{align*}
and $T^{(b)}_n = \sup_x|Y^{(b)}_n(x)|$.

\jy{Summary the procedure in an algorithm.}

The p-value of the block bootstrap KS test can be computed
\jy{not right}
as $p = \#\{T^{(b)}_n < T_n\} / B$ for 
$b \in \{1, \ldots, B\}$.

@@
